---
title: "Barriers and perceptions"
author: "Dennis van Soest"
output: 
  html_document: 
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(RColorBrewer)
data <- read.csv(file="//its-rds/2017/tightmr-transport-walking/Data_analysis2.csv", header=TRUE, sep=",")

```

##Hierarchical Clustering (data-driven)

Based on the attitudes (walking, cycling, driving, bus, train).

```{r echo=F}
dataset <- data[complete.cases(data[,c("Attit_walk", "Attit_Bus", "Attit_Train", "Attit_cycle", "Attit_car")]),]
clusters <- hclust(dist(dataset[,c("Attit_walk", "Attit_Bus", "Attit_Train", "Attit_cycle", "Attit_car")]), method="ward.D")
plot(clusters)
```


##Explore clusters
Characteristics of the three clusters.

```{r echo=F, warning=F, fig.width=5}
dataset$cluster <- as.factor(cutree(clusters, 4))

ggplot(dataset, aes(x=cluster, y=Attit_walk)) +
  geom_boxplot() +
  labs(title="Walk attitude", y = "Attitude")

ggplot(dataset, aes(x=cluster, y=Attit_Bus)) +
  geom_boxplot() +
  labs(title="Bus attitude", y = "Attitude")

ggplot(dataset, aes(x=cluster, y=Attit_Train)) +
  geom_boxplot() +
  labs(title="Train attitude", y = "Attitude")

ggplot(dataset, aes(x=cluster, y=Attit_car)) +
  geom_boxplot() +
  labs(title="Car attitude", y = "Attitude")

ggplot(dataset, aes(x=cluster, y=Attit_cycle)) +
  geom_boxplot() +
  labs(title="Cycling attitude", y = "Attitude")
```

##Barriers

```{r barrier plot preparation, echo=FALSE}
walk_barriers <- c("Bar_walk1","Bar_walk2","Bar_walk3","Bar_walk4","Bar_walk5",
                   "Bar_walk6","Bar_walk7","Bar_walk8","Bar_walk9", "Bar_walk10", "Bar_walk11")
bus_barriers <- c("Bar_bus1","Bar_bus2","Bar_bus3","Bar_bus4","Bar_bus5",
                  "Bar_bus6","Bar_bus7","Bar_bus8","Bar_bus9")
train_barriers <- c("Bar_train1","Bar_train2","Bar_train3","Bar_train4","Bar_train5",
                    "Bar_train6","Bar_train7","Bar_train8","Bar_train9")

descriptions_walk <- c("Walking takes too much time",
                  "I do not feel safe in traffic",
                  "I feel unsafe at night",
                  "I feel unsafe during the day",
                  "The walking routes are unpleasant",
                  "The pavements are bad",
                  "I am often too tired to walk",
                  "I am too lazy to walk",
                  "I won't walk when it is raining",
                  "I won't walk when I need to carry some luggage",
                  "I'd like to have more places to rest")
descriptions_bus <- c("The distance to the bus stop is too far",
                      "Travelling by bus takes too much time for my trips",
                      "I do not feel safe on buses",
                      "I won't take the bus if I need to carry some luggage",
                      "Buses are often too crowded for me",
                      "Buses are too expensive",
                      "The bus system is too complicated",
                      "I don't like the other people on the bus",
                      "I suffer from travel sickness on the bus")
descriptions_train <- c("The distance to the train station is too far",
                      "Travelling by train takes too much time for my trips",
                      "I do not feel safe on trains and at stations",
                      "I won't take the train if I need to carry some luggage",
                      "Trains are often too crowded for me",
                      "Trains are too expensive",
                      "The railway system is too complicated",
                      "I don't like the other people on the train",
                      "I suffer from travel sickness on the train")
clusters <- c("1", "2", "3", "4")

barrier_plots <- function(mode){
  barriers <- get(paste0(mode,"_barriers"))
  descriptions <- get(paste0("descriptions_", mode))
  for(i in 1:length(barriers)){
    plotdata <- na.omit(dataset[,c("cluster",barriers[[i]])])
    
    Agree_mat <- matrix(data=NA, nrow = 4*7, ncol = 3)
    colnames(Agree_mat) <- c("Cluster", "Rating", "Frequency")
    Rating <- c("Strongly disagree", "Disagree", "Somewhat disagree", "Neutral", "Somewhat agree", "Agree", "Strongly agree")
    for(n in 1:length(clusters)){
      for(r in 1:length(Rating)){
        Agree_mat[(n-1)*7+r,1]<- clusters[n]
        Agree_mat[(n-1)*7+r,2]<- Rating[r]
        Agree_mat[(n-1)*7+r,3]<- sum(plotdata[plotdata$cluster==clusters[n],2]==r)
      }
    } 
    Agree_mat.df <- as.data.frame(Agree_mat)
    Agree_mat.df$Frequency <- as.numeric(as.character(Agree_mat.df$Frequency))
    Agree_mat.df$Rating <- factor(Agree_mat.df$Rating, Rating)
    Agree_mat.df$Cluster <- as.factor(Agree_mat.df$Cluster)
    
    description <- paste0("Agreement with ",barriers[[i]])
    barchart <- ggplot(Agree_mat.df, aes(x=Cluster, y=Frequency, fill=Rating)) +
      geom_bar(position="fill", stat="identity", width = 0.5, col = "#424242") +
      scale_y_continuous(labels = percent_format()) +
      theme_grey() +
      coord_flip() +
      labs(y="Level of agreement", title=descriptions[[i]]) +
      #scale_x_discrete(limits = positions, labels=labels) + #order levels
      scale_fill_manual(values=brewer.pal(7, "Blues")) +
      guides(fill = guide_legend(keyheight = 0.5))
    print(barchart)
  }
}
```

###Barriers to walking

```{r walk barrier plots, fig.height=3, fig.width=7, echo=FALSE}
barrier_plots("walk")
```

###Barriers to bus use

```{r bus barrier plots, fig.height=3, fig.width=7, echo=FALSE}
barrier_plots("bus")
```

###Barriers to train use

```{r train barrier plots, fig.height=3, fig.width=7, echo=FALSE}
barrier_plots("train")
```

##Socio-demographic characteristics of the four clusters

Clusters per neighbourhood

```{r, echo=FALSE}
ggplot(dataset, aes(x=cluster, fill=Neighbourhood)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  theme(legend.position="bottom")+
  theme(legend.position="bottom",legend.title = element_text(size = 14) ,legend.text = element_text(size = 14))

ggplot(dataset, aes(x=Neighbourhood, fill=cluster)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  theme(legend.position="bottom",legend.title = element_text(size = 14) ,legend.text = element_text(size = 14))
```

Gender

```{r, echo=FALSE}
ggplot(na.omit(dataset[,c("cluster","Gender")]), aes(x=cluster, fill=as.factor(Gender))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(name="Gender", values=c("grey", "#00aeff", "#ff78eb"), labels=c("Not say", "Male", "Female"))
```

Age

```{r, echo=FALSE}
dataset$Year_birth[dataset$Year_birth==0] <- NA

ggplot(na.omit(dataset[,c("cluster","Year_birth")]), aes(x=Year_birth, fill=cluster)) +
  geom_density(alpha=0.4) + 
  labs(x="Year of birth")
```

Household size

```{r household size, echo=FALSE}
dataset$HH_size[dataset$HH_size==0] <- NA
dataset$HH_size <- as.factor(dataset$HH_size)
ggplot(na.omit(dataset[,c("cluster","HH_size")]), aes(x=cluster, fill=HH_size)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(7, "Blues"), name="Household size", 
                    labels=c("1 person", "2 people", "3 people", "4 people", "5 people", "6 people", "7 or more people"))
```

Children 

```{r household children, echo=FALSE}
dataset$HH_u12[dataset$HH_u12==9] <- NA
dataset$HH_u12 <- as.factor(dataset$HH_u12)
ggplot(na.omit(dataset[,c("cluster","HH_u12")]), aes(x=cluster, fill=HH_u12)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(4, "Blues"), name="Number of children in household", 
                    labels=c("None", "1 child", "2 children", "3 or more children"))
```

Income

```{r income, echo=FALSE}
dataset$Income[dataset$Income==0] <- NA

dataset$Income <- factor(dataset$Income, labels=c("< £15k","£15-25k","£25-50k","> £50k"))
ggplot(na.omit(dataset[,c("cluster","Income")]), aes(x=cluster, fill=Income)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(4, "Blues"), name="Annual household income")
```

Ethnicity

```{r ethnicity, echo=FALSE}
dataset$Ethnicity[dataset$Ethnicity==0] <- NA
dataset$Ethnicity <- factor(dataset$Ethnicity, labels = c("Asian/Asian British","Black/African/Caribbean/Black British",
                                           "Mixed/multiple ethnic group","White","Other ethnic group"))
ggplot(na.omit(dataset[,c("cluster","Ethnicity")]), aes(x=cluster, fill=Ethnicity)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Set1"))
```

Work status

```{r work status, echo=FALSE}
dataset$Work_status[dataset$Work_status==0] <- NA

dataset$Work_status <- factor(dataset$Work_status, labels=c("Employed full time", "Employed part time", "Student", "Retired", "Unemployed"))
ggplot(na.omit(dataset[,c("cluster","Work_status")]), aes(x=cluster, fill=Work_status)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Set1"), name="Work status")
```

##Travel characteristics of the three clusters

Use of walking

```{r, echo=FALSE}
ggplot(na.omit(dataset[,c("cluster","Use_walk")]), aes(x=cluster, fill=as.factor(Use_walk))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Blues"), name="Walk frequency", 
                    labels=c("Never", "<1 day per week", "1 or 2 days per week", "3 to 5 days per week", "6 or 7 days per week"))
```

Use of car as a driver

```{r, echo=FALSE}
ggplot(na.omit(dataset[,c("cluster","Use_drive")]), aes(x=cluster, fill=as.factor(Use_drive))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Blues"), name="Driving frequency", 
                    labels=c("Never", "<1 day per week", "1 or 2 days per week", "3 to 5 days per week", "6 or 7 days per week"))
```

Use of car as a passenger

```{r, echo=FALSE}
ggplot(na.omit(dataset[,c("cluster","Use_pass")]), aes(x=cluster, fill=as.factor(Use_pass))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Blues"), name="Car passenger frequency", 
                    labels=c("Never", "<1 day per week", "1 or 2 days per week", "3 to 5 days per week", "6 or 7 days per week"))
```

Use of train

```{r, echo=FALSE}
ggplot(na.omit(dataset[,c("cluster","Use_train")]), aes(x=cluster, fill=as.factor(Use_train))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Blues"), name="Train use frequency", 
                    labels=c("Never", "<1 day per week", "1 or 2 days per week", "3 to 5 days per week", "6 or 7 days per week"))
```

Use of bus

```{r, echo=FALSE}
ggplot(na.omit(dataset[,c("cluster","Use_bus")]), aes(x=cluster, fill=as.factor(Use_bus))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Blues"), name="Bus use frequency", 
                    labels=c("Never", "<1 day per week", "1 or 2 days per week", "3 to 5 days per week", "6 or 7 days per week"))
```

Use of bicycle

```{r, echo=FALSE}
ggplot(na.omit(dataset[,c("cluster","Use_cycling")]), aes(x=cluster, fill=as.factor(Use_cycling))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Blues"), name="Cycling frequency", 
                    labels=c("Never", "<1 day per week", "1 or 2 days per week", "3 to 5 days per week", "6 or 7 days per week"))
```

Use of taxi

```{r, echo=FALSE}
ggplot(na.omit(dataset[,c("cluster","Use_taxi")]), aes(x=cluster, fill=as.factor(Use_taxi))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values=brewer.pal(5, "Blues"), name="Taxi use frequency", 
                    labels=c("Never", "<1 day per week", "1 or 2 days per week", "3 to 5 days per week", "6 or 7 days per week"))
```


##Associations

```{r}
as_data <- read.csv(file="//its-rds/2017/tightmr-transport-walking/AssociationData/AssociationData3_csv.csv", header=TRUE, sep=",")
names(dataset)[1] <- "Respondent"
com_data <- merge(x=as_data, y=dataset, by="Respondent")
com_data$Code <- as.factor(com_data$Code)
com_data$cluster <- as.factor(com_data$cluster)
modes <- list("walk", "car", "train", "cycle", "bus")

source("C:/Users/dxv632/OneDrive - University of Birmingham/_PhD/_Parts/Questionnaire/AssociationAnalysis/Analysis_function_moonplot.R")

analysis_moon(v="cluster")
```


